---
author: "Hanna Lee"
date: "2/22/2022"
title: "Incarceration Data Visualization"
output: html_document
---

```{r, setup, include = FALSE}
library("tidyverse")
library("knitr")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "../source")
```

```{r, sources, include = FALSE}
source("~/Documents/INFO_201/a3-hlee107/source/analysis.R")
source("~/Documents/INFO_201/a3-hlee107/source/a3-functions.R")
```

# Report {.tabset .tabset-fade .tabset-pills}

### Introduction
Incarceration is a multifaceted issue. From this dataset, I hope to extrapolate information from the story behind the numbers. My focus is on Washington state.
I will evaluate the trends in the incarceration population in the BIPOC 
community for the current top 5 populous counties in Washington. Since
Washington is labeled as a liberal state, I wanted to compare it to the most
conservative state in the country, Mississippi. I first evaluated the jail
population of Black people in Mississippi and Washington. Then, I wanted to see
a further comparison with the incarceration population of Black and White people
in Washington and Mississippi. These charts and maps will help us summarise the
relationship between an inmate's race and location.

<br>

### Summary of Information

<br>

## Chart that Shows Trends Over Time

### <u>County and BIPOC Jail Population</u>

<br>

```{r, over_time, echo = FALSE, fig.show="hold", out.width="100%"}
data <- population_data(filter_jail_pop)
plot(bipoc_pop_line(data))
```
<br>
<br>
<div style="background-color:#F2EEED;padding:5px;">
<p style="text-indent:40px;font-size:17px;">
Both graphs show a slightly negatively sloped linear regression line, which would indicate that the higher the population proportion with at least two access points (relative to one), the less students meet test standards. This can loosely imply that more Internet access points does not result in better education. However, correlation does not imply causation, and so further tests would be required. Moreover, if you remove some points, the graph appears it could have a positively sloped regression line.
</p></div>
<!-------------------------------------------------------------------------------------------------->

## Chart that Compares Two Variables

### <u>State and Jail Population of Black People</u>

<br>

```{r, compare, echo = FALSE, fig.show="hold", out.width="100%"}
data_2 <- ms_wa_black_pop
plot(state_black_scatter(data_2))
```


## Map of Washington State and Mississippi

### <u>State and Jail Population of Black and White People</u>

<br>

```{r, map, echo = FALSE, fig.show="hold", out.width="50%"}
data_3 <- wa_map_data
data_4 <- ms_map_data
wa_black_map(data_3)
wa_white_map(data_3)
ms_black_map(data_4)
ms_white_map(data_4)
```
