---
author: "Hanna Lee"
date: "2/22/2022"
title: "Incarceration Data Visualization"
output: html_document
---

```{r, setup, include = FALSE}
library("tidyverse")
library("knitr")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "../source")
```

```{r, sources, include = FALSE}
source("~/Documents/INFO_201/a3-hlee107/source/analysis.R")
source("~/Documents/INFO_201/a3-hlee107/source/a3-functions.R")
```

# Report {.tabset .tabset-fade .tabset-pills}

### Introduction

<br>

### Summary of Information

<br>

## Chart that Shows Trends Over Time

### <u>County and BIPOC Jail Population</u>

<br>

```{r, over_time, echo = FALSE, fig.show="hold", out.width="100%"}
data <- population_data(filter_jail_pop)
plot(bipoc_pop_line(data))
```
<br>
<br>
<div style="background-color:#F2EEED;padding:5px;">
<p style="text-indent:40px;font-size:17px;">
Both graphs show a slightly negatively sloped linear regression line, which would indicate that the higher the population proportion with at least two access points (relative to one), the less students meet test standards. This can loosely imply that more Internet access points does not result in better education. However, correlation does not imply causation, and so further tests would be required. Moreover, if you remove some points, the graph appears it could have a positively sloped regression line.
</p></div>
<!-------------------------------------------------------------------------------------------------->

## Chart that Compares Two Variables

### <u>State and Jail Population of Black People</u>

<br>

```{r, compare, echo = FALSE, fig.show="hold", out.width="100%"}
data_2 <- ms_wa_black_pop
plot(state_black_scatter(data_2))
```


## Map of Washington State and Mississippi

### <u>State and Jail Population of Black and White People</u>

<br>

```{r, map, echo = FALSE, fig.show="hold", out.width="50%"}
data_3 <- wa_map_data
data_4 <- ms_map_data
wa_black_map(data_3)
wa_white_map(data_3)
ms_black_map(data_4)
ms_white_map(data_4)
```
